---
title: |
  MAKING A MOVIE <br> WHO SHOULD WE HIRE 🧐
author: "Tam, Nauman, Sushant, Ahmad, Irtasam"
date: "`r Sys.Date()`"
fontsize: 11pt
output:
  rmdformats::readthedown:
  number_sections: true
---

```{r setup, include=FALSE, cache=FALSE}
## Libraries
## You may have to install these libraries first 
## with the tab on bottom right panel : "Packages" ; click install
## or : install.packages("ggplot2")   - don't forget the ""
library(knitr)
library(rmdformats)

## Global options
knitr::opts_chunk$set(               # On définit ici les comportements par défaut des capsules dans le document produit html
               echo=T,        # pour afficher les lignes de code (FALSE pour les cacher ; mais elles seront executer)
               eval=TRUE,     # pour executer les capsules de code
	             message=FALSE, # pour que les messages de R ne soient pas affichés dans le doc html
               warning=FALSE) # pour que les alertes de R ne soient pas affichés dans le doc html
knitr::opts_knit$set(width=100)  
options(max.print="100")       # pour définir le nombre d'éléments afficher par R

options(max.print="100")
opts_chunk$set(echo=T,    # pour afficher les lignes de code (FALSE pour les cacher ; mais elles seront executer)  
               eval=TRUE, # pour executer les capsules de code
               message=FALSE,  # pour que les messages de R ne soient pas affichés dans le doc html
               warning=FALSE)  # pour que les alertes de R ne soient pas affichés dans le doc html
opts_knit$set(width=100)  # pour définir le nombre d'éléments afficher par R
```

<!-- Enoncé -->

In this data analysis project, we delve into the movie dataset sourced from I. Joly & P. Lemaire. Our approach follows the CRISP-DM methodology. The results obtained provide valuable insights, shaping decisions and strategies in the realm of filmmaking.

# I. Business understanding {.tabset}

Talk about the project what kind of topic 

# II. Data understanding {.tabset}

Then we talk about the params the param we gonna to use with reason - guess
```{r, echo=FALSE}
# Use for display the main params we going to use in this project
params <- c("movie_title", "actor_1_name", "imdb_score", "genres", "ROI_perc")
descriptions <- c(
  "Title of the movie",
  "Primary actor starring in the movie",
  "IMDB Score of the movie on IMDB, rated by the users",
  "Film categorization",
  "ROI = profit / budget"
)

# Create a data frame with a column for numbering
table_data <- data.frame(
  No. = 1:length(params),
  Parameter = params,
  Description = descriptions
)

# Print the table using kable
knitr::kable(table_data, col.names = c("No.", "Parameter", "Description"), align = "l")

```

# III. Data preparation {.tabset}

## General Movie.Data
```{r, echo=FALSE}
# List of required libraries
required_libraries <- c("rmdformats", "skimr", "readr", "dplyr", "anytime", "gt", "magrittr", "ggplot2", "stringr")

# Set a CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Check if each library is installed, and install if not
for (lib in required_libraries) {
  if (!requireNamespace(lib, quietly = TRUE)) {
    install.packages(lib, dependencies = TRUE)
  }
  library(lib, character.only = TRUE)
}
```

Import these libraries to run this R.script of this project
```{r, echo=FALSE}
library(rmdformats)
library(skimr)
library(readr)
library(dplyr)
library(anytime)
library(gt)
library(magrittr)
library(ggplot2)
library(stringr)
```

```{r, echo=FALSE}
# List of libraries
library_list <- c(
  "rmdformats" = "Custom R Markdown formats and templates",
  "skimr" = "Summarizes data frames and vectors",
  "readr" = "Reads flat files (e.g., CSV, TSV)",
  "dplyr" = "Data manipulation and transformation",
  "anytime" = "Converts dates and times",
  "gt" = "Generates tables with the gt package",
  "magrittr" = "Provides the pipe operator %>%",
  "ggplot2" = "Data visualization with ggplot2",
  "stringr" = "String manipulation functions"
)

# Create a data frame
lib_data <- data.frame(
  No = 1:length(library_list),
  Name = names(library_list),
  Description = unname(library_list)
)

# Display the table using kable
knitr::kable(lib_data, col.names = c("No.", "Lib Name", "Description"), align = "l")
```

Data loading
```{r}
movies_data <- load("Movie.Data_Complete.RData")
```

```{r, echo=FALSE}
Movie.Data %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("General View of the Dataset")
```

Displaying summary statistics
```{r}
Movie.Data %>% summary() %>% kable() %>% head()
```

## Specific data for project

Because of the ... we only take the most frequency genres in the dataset
```{r, echo=FALSE}
# Extract individual genres from each row
all_genres <- str_split(Movie.Data$genres, "\\|") %>%
  unlist()

# Create a data frame with the counts of each genre
genre_counts <- data.frame(genre = all_genres) %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  arrange(desc(count))  # Arrange in descending order

# Select the top 10 genres
top_10_genres <- head(genre_counts, 10)

# Add a numbering column
top_10_genres$No. <- seq_len(nrow(top_10_genres))

# Display the top 10 genres in a table
knitr::kable(top_10_genres[, c("No.", "genre", "count")], col.names = c("No.", "Genres", "Count"), align = "l")
```

Create dataset for the project called "selected_genres_subset"
```{r, echo=FALSE}
# Specify the genres you are interested in
selected_genres <- c('Drama', 'Comedy', 'Thriller', 'Action', 'Romance')

# Create a subset of the dataset with specific genres and columns
selected_genres_subset <- Movie.Data %>%
  filter(grepl(paste(selected_genres, collapse = '|'), genres, ignore.case = TRUE)) %>%
  select(
    movie_title, actor_1_name, genres, ROI_perc, imdb_score
  )

# Display the first few rows of the subset using gt
selected_genres_subset %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_header(
    title = "Subset of Movies with Selected Genres and Specific Columns"
  ) %>%
  tab_caption("Selected Genres Subset")

# Display the summary
selected_genres_subset %>% summary() %>% kable() %>% head()
```

Check for missing values in the data
```{r, echo=FALSE}
missing_values <- sapply(selected_genres_subset, function(x) sum(is.na(x)))
# Display the number of missing values for each column
print(missing_values)
```

We going to remove 994 NA value ROI (because we also check about the profit it have same NA value so that we couldn't cal the ROI) 
```{r, echo=FALSE}
# Remove rows with missing values in 'ROI_perc'
selected_genres_subset <- selected_genres_subset %>%
  filter(!is.na(ROI_perc))

# Display the summary
selected_genres_subset %>% summary() %>% kable() %>% head()

# Extract individual genres from each row
all_genres <- str_split(selected_genres_subset$genres, "\\|") %>%
  unlist()

# Create a data frame with the counts of each genre
genre_counts <- data.frame(genre = all_genres) %>%
  group_by(genre) %>%
  summarise(count = n()) %>%
  arrange(desc(count))  # Arrange in descending order

# Select the top 5 genres
top_5_genres <- head(genre_counts, 5)

# Add a numbering column
top_5_genres$No. <- seq_len(nrow(top_5_genres))

# Display the top 5 genres in a table
knitr::kable(top_5_genres[, c("No.", "genre", "count")], col.names = c("No.", "Genres", "Count"), align = "l")
```

Now we separate into 5 small table 
```{r, echo=FALSE}
# Initialize a list to store the small tables
small_tables_list <- list()

# Initialize a list to store HTML tables
html_tables_list <- list()

# Iterate over each genre and create a small table
for (genre in selected_genres) {
  # Filter rows for the current genre
  current_genre_table <- selected_genres_subset %>%
    filter(grepl(genre, genres, ignore.case = TRUE)) %>%
    select(
      movie_title, actor_1_name, genres, ROI_perc, imdb_score
    )
  
  # Save the small table into the list
  small_tables_list[[paste0(genre, "_table")]] <- current_genre_table
  
  # Convert the small table to HTML format
  html_table <- knitr::kable(current_genre_table, col.names = c("Movie Title", "Actor 1 Name", "Genres", "ROI Perc", "IMDB Score"), align = "l")
  
  # Save the HTML table into the list
  html_tables_list[[paste0(genre, "_html_table")]] <- html_table
  
  # Print the first few rows of the small table
#  cat("\n", genre, "Table:\n")
# print(head(current_genre_table))
}

# Save all small tables into the R environment
list2env(small_tables_list, envir = .GlobalEnv)
```
```{r, echo=FALSE}
# Display the first few rows of the subset
Drama_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Drama Genres Subset")
Drama_table %>% summary() %>% kable() %>% head()

Comedy_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Comedy Genres Subset")
Comedy_table %>% summary() %>% kable() %>% head()

Thriller_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Thriller Genres Subset")
Thriller_table %>% summary() %>% kable() %>% head()

Action_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Action Genres Subset")
Action_table %>% summary() %>% kable() %>% head()

Romance_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Romance Genres Subset")
Romance_table %>% summary() %>% kable() %>% head()
```


# IV. Modeling based on imdb_score {.tabset}

## Drama Table

Put chart and comment of that chart

Check in each genres_table that if actor_1_name appear less than 5 times remove that line out of table

```{r, echo=FALSE}
# Check the count of each actor's appearances in 'Drama_table'
actor_counts <- Drama_table %>%
  group_by(actor_1_name) %>%
  summarise(count = n())

# Filter out actors with appearances less than 5
actors_to_remove <- actor_counts %>%
  filter(count < 5) %>%
  pull(actor_1_name)

# Remove rows with actors to be removed from 'Drama_table'
Drama_table_filtered <- Drama_table %>%
  filter(!actor_1_name %in% actors_to_remove)

# Display the result using gt
Drama_table_filtered %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Filtered Action Table")

Drama_table_filtered %>% summary() %>% kable() %>% head()

# Check the total count of unique actor_1_name in 'Drama_table_filtered'
total_unique_actors <- Drama_table_filtered %>%
  summarise(total_actors = n_distinct(actor_1_name))
# Display the result as text
cat("But in 724 rows, it actually only", total_unique_actors$total_actors, "actors.")
```

Cal the actor imdb 
```{r, echo=FALSE}
# Calculate the average imdb_score for each actor appearing more than 5 times
actor_avg_scores <- Drama_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(imdb_actor = mean(imdb_score, na.rm = TRUE))

# Merge the actor_avg_scores back into 'Drama_table_filtered'
Drama_table_with_avg <- left_join(Drama_table_filtered, actor_avg_scores, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, imdb_actor, and ranking
actor_ranking_table <- actor_avg_scores %>%
  arrange(desc(imdb_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, imdb_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table
actor_ranking_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table, aes(x = actor_1_name, y = imdb_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", imdb_actor), y = imdb_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor IMDb Scores and Rankings on Drama genres",
    x = "Actor Name",
    y = "Average IMDb Score",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(6, 8)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Comedy Table

Put chart and comment of that chart

Check in each genres_table that if actor_1_name appear less than 5 times remove that line out of table

```{r, echo=FALSE}
# Check the count of each actor's appearances in 'Comedy_table'
actor_counts <- Comedy_table %>%
  group_by(actor_1_name) %>%
  summarise(count = n())

# Filter out actors with appearances less than 5
actors_to_remove <- actor_counts %>%
  filter(count < 5) %>%
  pull(actor_1_name)

# Remove rows with actors to be removed from 'Comedy_table'
Comedy_table_filtered <- Comedy_table %>%
  filter(!actor_1_name %in% actors_to_remove)

# Display the result using gt
Comedy_table_filtered %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Filtered Action Table")

Comedy_table_filtered %>% summary() %>% kable() %>% head()

# Check the total count of unique actor_1_name in 'Comedy_table_filtered'
total_unique_actors <- Comedy_table_filtered %>%
  summarise(total_actors = n_distinct(actor_1_name))
# Display the result as text
cat("But in 526 rows, it actually only", total_unique_actors$total_actors, "actors.")
```

Cal the actor imdb 
```{r, echo=FALSE}
# Calculate the average imdb_score for each actor appearing more than 5 times
actor_avg_scores <- Comedy_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(imdb_actor = mean(imdb_score, na.rm = TRUE))

# Merge the actor_avg_scores back into 'Comedy_table_filtered'
Comedy_table_with_avg <- left_join(Comedy_table_filtered, actor_avg_scores, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, imdb_actor, and ranking
actor_ranking_table <- actor_avg_scores %>%
  arrange(desc(imdb_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, imdb_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table
actor_ranking_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table, aes(x = actor_1_name, y = imdb_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", imdb_actor), y = imdb_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor IMDb Scores and Rankings on Comedy genres",
    x = "Actor Name",
    y = "Average IMDb Score",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(6, 8)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Thriller Table

Put chart and comment of that chart

Check in each genres_table that if actor_1_name appear less than 5 times remove that line out of table

```{r, echo=FALSE}
# Check the count of each actor's appearances in 'Thriller_table'
actor_counts <- Thriller_table %>%
  group_by(actor_1_name) %>%
  summarise(count = n())

# Filter out actors with appearances less than 5
actors_to_remove <- actor_counts %>%
  filter(count < 5) %>%
  pull(actor_1_name)

# Remove rows with actors to be removed from 'Thriller_table'
Thriller_table_filtered <- Thriller_table %>%
  filter(!actor_1_name %in% actors_to_remove)

# Display the result using gt
Thriller_table_filtered %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Filtered Action Table")

Thriller_table_filtered %>% summary() %>% kable() %>% head()

# Check the total count of unique actor_1_name in 'Thriller_table_filtered'
total_unique_actors <- Thriller_table_filtered %>%
  summarise(total_actors = n_distinct(actor_1_name))
# Display the result as text
cat("But in 371 rows, it actually only", total_unique_actors$total_actors, "actors.")
```

Cal the actor imdb 
```{r, echo=FALSE}
# Calculate the average imdb_score for each actor appearing more than 5 times
actor_avg_scores <- Thriller_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(imdb_actor = mean(imdb_score, na.rm = TRUE))

# Merge the actor_avg_scores back into 'Thriller_table_filtered'
Thriller_table_with_avg <- left_join(Thriller_table_filtered, actor_avg_scores, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, imdb_actor, and ranking
actor_ranking_table <- actor_avg_scores %>%
  arrange(desc(imdb_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, imdb_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table
actor_ranking_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table, aes(x = actor_1_name, y = imdb_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", imdb_actor), y = imdb_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor IMDb Scores and Rankings on Thriller genres",
    x = "Actor Name",
    y = "Average IMDb Score",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(6, 8)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Action Table

Put chart and comment of that chart

Check in each genres_table that if actor_1_name appear less than 5 times remove that line out of table

```{r, echo=FALSE}
# Check the count of each actor's appearances in 'Action_table'
actor_counts <- Action_table %>%
  group_by(actor_1_name) %>%
  summarise(count = n())

# Filter out actors with appearances less than 5
actors_to_remove <- actor_counts %>%
  filter(count < 5) %>%
  pull(actor_1_name)

# Remove rows with actors to be removed from 'Action_table'
Action_table_filtered <- Action_table %>%
  filter(!actor_1_name %in% actors_to_remove)

# Display the result using gt
Action_table_filtered %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Filtered Action Table")

Action_table_filtered %>% summary() %>% kable() %>% head()

# Check the total count of unique actor_1_name in 'Action_table_filtered'
total_unique_actors <- Action_table_filtered %>%
  summarise(total_actors = n_distinct(actor_1_name))
# Display the result as text
cat("But in 335 rows, it actually only", total_unique_actors$total_actors, "actors.")
```

Cal the actor imdb 
```{r, echo=FALSE}
# Calculate the average imdb_score for each actor appearing more than 5 times
actor_avg_scores <- Action_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(imdb_actor = mean(imdb_score, na.rm = TRUE))

# Merge the actor_avg_scores back into 'Action_table_filtered'
Action_table_with_avg <- left_join(Action_table_filtered, actor_avg_scores, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, imdb_actor, and ranking
actor_ranking_table <- actor_avg_scores %>%
  arrange(desc(imdb_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, imdb_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table
actor_ranking_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table, aes(x = actor_1_name, y = imdb_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", imdb_actor), y = imdb_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor IMDb Scores and Rankings on Action Genres",
    x = "Actor Name",
    y = "Average IMDb Score",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(6, 8)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Romance Table

Put chart and comment of that chart

Check in each genres_table that if actor_1_name appear less than 5 times remove that line out of table

```{r, echo=FALSE}
# Check the count of each actor's appearances in 'Romance_table'
actor_counts <- Romance_table %>%
  group_by(actor_1_name) %>%
  summarise(count = n())

# Filter out actors with appearances less than 5
actors_to_remove <- actor_counts %>%
  filter(count < 5) %>%
  pull(actor_1_name)

# Remove rows with actors to be removed from 'Drama_table'
Romance_table_filtered <- Romance_table %>%
  filter(!actor_1_name %in% actors_to_remove)

# Display the result using gt
Romance_table_filtered %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Filtered Action Table")

Romance_table_filtered %>% summary() %>% kable() %>% head()

# Check the total count of unique actor_1_name in 'Romance_table_filtered'
total_unique_actors <- Romance_table_filtered %>%
  summarise(total_actors = n_distinct(actor_1_name))
# Display the result as text
cat("But in 175 rows, it actually only", total_unique_actors$total_actors, "actors.")
```

Cal the actor imdb 
```{r, echo=FALSE}
# Calculate the average imdb_score for each actor appearing more than 5 times
actor_avg_scores <- Romance_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(imdb_actor = mean(imdb_score, na.rm = TRUE))

# Merge the actor_avg_scores back into 'Romance_table_filtered'
Romance_table_with_avg <- left_join(Romance_table_filtered, actor_avg_scores, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, imdb_actor, and ranking
actor_ranking_table <- actor_avg_scores %>%
  arrange(desc(imdb_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, imdb_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table
actor_ranking_table %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table, aes(x = actor_1_name, y = imdb_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", imdb_actor), y = imdb_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor IMDb Scores and Rankings on Romance genres",
    x = "Actor Name",
    y = "Average IMDb Score",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(6, 8)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```


# V. Modeling based on ROI_perc {.tabset}

## Drama Table
Cal the actor ROI 
```{r, echo=FALSE}
# Calculate the average ROI_perc for each actor appearing more than 5 times
actor_avg_roi <- Drama_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(roi_actor = mean(ROI_perc, na.rm = TRUE))

# Merge the actor_avg_roi back into 'Drama_table_filtered'
Drama_table_with_avg_roi <- left_join(Drama_table_filtered, actor_avg_roi, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, roi_actor, and ranking
actor_ranking_table_roi <- actor_avg_roi %>%
  arrange(desc(roi_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, roi_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table_roi
actor_ranking_table_roi %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table_roi, aes(x = actor_1_name, y = roi_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", roi_actor), y = roi_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor ROI_perc and Rankings on Drama genres",
    x = "Actor Name",
    y = "Average ROI_perc",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(300, 1800)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Comedy Table

Cal the actor ROI 
```{r, echo=FALSE}
# Calculate the average ROI_perc for each actor appearing more than 5 times
actor_avg_roi <- Comedy_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(roi_actor = mean(ROI_perc, na.rm = TRUE))

# Merge the actor_avg_roi back into 'Comedy_table_filtered'
Comedy_table_with_avg_roi <- left_join(Comedy_table_filtered, actor_avg_roi, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, roi_actor, and ranking
actor_ranking_table_roi <- actor_avg_roi %>%
  arrange(desc(roi_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, roi_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table_roi
actor_ranking_table_roi %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table_roi, aes(x = actor_1_name, y = roi_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", roi_actor), y = roi_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor ROI_perc and Rankings on Comedy genres",
    x = "Actor Name",
    y = "Average ROI_perc",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(200, 850)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Thriller Table

Cal the actor ROI 
```{r, echo=FALSE}
# Calculate the average ROI_perc for each actor appearing more than 5 times
actor_avg_roi <- Thriller_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(roi_actor = mean(ROI_perc, na.rm = TRUE))

# Merge the actor_avg_roi back into 'Thriller_table_filtered'
Thriller_table_with_avg_roi <- left_join(Thriller_table_filtered, actor_avg_roi, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, roi_actor, and ranking
actor_ranking_table_roi <- actor_avg_roi %>%
  arrange(desc(roi_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, roi_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table_roi
actor_ranking_table_roi %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table_roi, aes(x = actor_1_name, y = roi_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", roi_actor), y = roi_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor ROI_perc and Rankings on Thriller genres",
    x = "Actor Name",
    y = "Average ROI_perc",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(50, 300)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Action Table
Cal the actor ROI 
```{r, echo=FALSE}
# Calculate the average ROI_perc for each actor appearing more than 5 times
actor_avg_roi <- Action_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(roi_actor = mean(ROI_perc, na.rm = TRUE))

# Merge the actor_avg_roi back into 'Action_table_filtered'
Action_table_with_avg_roi <- left_join(Action_table_filtered, actor_avg_roi, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, roi_actor, and ranking
actor_ranking_table_roi <- actor_avg_roi %>%
  arrange(desc(roi_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, roi_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table_roi
actor_ranking_table_roi %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table_roi, aes(x = actor_1_name, y = roi_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", roi_actor), y = roi_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor ROI_perc and Rankings on Action genres",
    x = "Actor Name",
    y = "Average ROI_perc",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(50, 650)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

## Romance Table
Cal the actor ROI 
```{r, echo=FALSE}
# Calculate the average ROI_perc for each actor appearing more than 5 times
actor_avg_roi <- Romance_table_filtered %>%
  group_by(actor_1_name) %>%
  summarise(roi_actor = mean(ROI_perc, na.rm = TRUE))

# Merge the actor_avg_roi back into 'Romance_table_filtered'
Romance_table_with_avg_roi <- left_join(Romance_table_filtered, actor_avg_roi, by = "actor_1_name")

# Display the first few rows of the merged table
#head(Action_table_with_avg)

# Create a table with actor names, roi_actor, and ranking
actor_ranking_table_roi <- actor_avg_roi %>%
  arrange(desc(roi_actor)) %>%
  mutate(ranking = row_number()) %>%
  select(actor_1_name, roi_actor, ranking) %>%
  filter(ranking <= 10)

# Display the actor_ranking_table_roi
actor_ranking_table_roi %>%
  head() %>%
  gt() %>%
  tab_spanner(label = "Variable") %>%
  tab_caption("Actor ranking Table")
```

Create bar chart
```{r}
# Create a bar chart with scaled y-axis and rotated x-axis labels
bar_chart_scaled <- ggplot(actor_ranking_table_roi, aes(x = actor_1_name, y = roi_actor, fill = factor(ranking))) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", roi_actor), y = roi_actor), vjust = -0.5) +  # Add value labels
  scale_fill_viridis_d() +
  labs(
    title = "Actor ROI_perc and Rankings on Romance genres",
    x = "Actor Name",
    y = "Average ROI_perc",
    fill = "Ranking"
  ) +
  theme_minimal() +
  coord_cartesian(ylim = c(100, 350)) +  # Set y-axis limit
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

# Display the scaled bar chart
print(bar_chart_scaled)
```

# VI. Evaluation {.tabset}

Evaluate results
Review process

# VII. Deployment {.tabset}
 
Final Report
